In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-30 paper/main.tex
Remove line breaks in abstract for convenient copy-pasting.
Add Rami acknowledgement of support from EMIx.
Fix figure references

We study the effect of such rapid PVS flow on intracranial molecular
enrichment by comparing the more conservative baseline model and a
high PVS flow model, where the former still includes the net flow
contributions from CSF production and the cardiac pulse wave, while
the latter additionally includes the ultraslow vasomotion
contribution. In both models, tracers were first observed at the basal
artery after 48 minutes (\Cref{fig:pvs_b}A), defined by their first-time arrival (FTA), the time at which the concentration first exceeds $0.1$ mmol/l. At all upstream locations along the middle cerebral arteries
(MCAs) and anterior cerebral artery (ACA), we found substantially
reduced FTAs with higher PVS flow (\Cref{fig:pvs_a}B), up to 2.5 hours
earlier in the left M2 segment of the MCA. PVS concentrations peaked
before the surrounding tissue (0:48h vs 1:12h at the MCA and 1:00h vs
1:48h at MCA2) with high PVS flow, while these peaks occurred nearly
simultaneously in the baseline model. The earlier appearance and the
delay in time-to-peak between the PVS and surrounding tissue clearly
indicates the directionality of tracer enrichment -- it first arrives
in the PVS and subsequently spreads into the tissue and SAS
(\Cref{fig:pvs_a}C, D), and especially so with higher PVS flow. These
observations also hold on an aggregated level: computing the total
amount of tracer in the PVS as a function of the distance to the
arterial network roots, we find accelerated tracer transport with
higher PVS flow, especially at earlier time points (2--9 hours)
(\Cref{fig:pvs_b}E). Moreover, differences in tracer transport along the
PVS translate to altered enrichment patterns on the whole-organ
scale. For instance, after 4-6 hours, the faster-moving tracer in the
PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs_b}G). In contrast,
there are no clear signs of early enrichment surrounding the PVS in
the baseline model (\Cref{fig:pvs_b}F).

@@

All authors express their gratitude to Prof.~Jürgen Reichenbach for
generously sharing the multi-modal human MR data, and to Prof. Erlend
Hodneland for facilitating this data exchange. M.E.R.~acknowledges
support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
Centre for Brain Fluid Research. M.E.R.~and M.C.~acknowledge support
from the Research Council of Norway (RCN) via grant \#324239
(EMIx). Simulations were performed on the Experimental Infrastructure
for Exploration of Exascale Computing (eX3), which is financially
supported by the RCN under contract 270053. M.K.~acknowledges support
from the RCN grant \#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A
were created in Biorender.

>>>>

We study the effect of such rapid PVS flow on intracranial molecular
enrichment by comparing the more conservative baseline model and a
high PVS flow model, where the former still includes the net flow
contributions from CSF production and the cardiac pulse wave, while
the latter additionally includes the ultraslow vasomotion
contribution. In both models, tracers were first observed at the basal
artery after 48 minutes (\Cref{fig:pvs_b}A), defined by their first-time arrival (FTA), the time at which the concentration first exceeds $0.1$ mmol/l. At all upstream locations along the middle cerebral arteries
(MCAs) and anterior cerebral artery (ACA), we found substantially
reduced FTAs with higher PVS flow (\Cref{fig:pvs_b}B), up to 2.5 hours
earlier in the left M2 segment of the MCA. PVS concentrations peaked
before the surrounding tissue (0:48h vs 1:12h at the MCA and 1:00h vs
1:48h at MCA2) with high PVS flow, while these peaks occurred nearly
simultaneously in the baseline model. The earlier appearance and the
delay in time-to-peak between the PVS and surrounding tissue clearly
indicates the directionality of tracer enrichment -- it first arrives
in the PVS and subsequently spreads into the tissue and SAS
(\Cref{fig:pvs_b}C, D), and especially so with higher PVS flow. These
observations also hold on an aggregated level: computing the total
amount of tracer in the PVS as a function of the distance to the
arterial network roots, we find accelerated tracer transport with
higher PVS flow, especially at earlier time points (2--9 hours)
(\Cref{fig:pvs_b}E). Moreover, differences in tracer transport along the
PVS translate to altered enrichment patterns on the whole-organ
scale. For instance, after 4-6 hours, the faster-moving tracer in the
PVS is clearly visible in the space adjacent to the PVS of the ACA, MCA, and other arteries (\Cref{fig:pvs_b}G). In contrast,
there are no clear signs of early enrichment surrounding the PVS in
the baseline model (\Cref{fig:pvs_b}F).

@@

All authors express their gratitude to Prof.~Jürgen Reichenbach for
generously sharing the multi-modal human MR data, and to Prof. Erlend
Hodneland for facilitating this data exchange. M.E.R.~acknowledges
support from Stiftelsen Kristian Gerhard Jebsen via the K.~G.~Jebsen
Centre for Brain Fluid Research. M.C., R.M., and M.E.R.~acknowledge
support from the Research Council of Norway (RCN) via grant \#324239
(EMIx). Simulations were performed on the Experimental Infrastructure
for Exploration of Exascale Computing (eX3), which is financially
supported by the RCN under contract 270053. M.K.~acknowledges support
from the RCN grant \#303362. The figures 1A, 1D, 2C, 3C, 5A and 6A
were created in Biorender.
